---
title: "Climate Legislation Statistics"
output: pdf_document
---

# Preprocessing

## Functions

First, some libraries are imported and functions made.
In addition to the functions in `functions.R`, the following are made:

* `annual_fish` - decrepit function that compares different types of legislation
* `manipulate` - preprocessing function that combines same names and removes JRP
* `com_comparison`, `com_comparison_readable` - Compares everything in a list of committees in a dataframe to everything else in dataframe.

In theory, should be the same, but has a very small and negligible difference in output

```{r Libraries and Functions, echo=FALSE}
options(warn = -1)

# Performs Chi-Squared tests on the different outcomes
source("functions.R")
library(stringr)

annual_fish <- function(climate, education, year = 2015) {
    cl <- filter(climate, Year == year)
    cl <- filter(cl, Disposition != "JRP") %>%
        mutate(Disposition = if_else(Disposition %in% c("DIC", "PIL"), Disposition, "DIO"))
    cl <- data.frame(dis = cl$Disposition, type = "Climate")
    e <- filter(education, Year == year) %>%
        mutate(Disposition = if_else(Disposition %in% c("DIC", "PIL"), Disposition, "DIO"))
    e <- data.frame(dis = e$Disposition, type = "Education")
    comb <- table(rbind(cl, e))
    print(comb)
    print(chisq.test(comb))
    return(comb)
}

sub <- function(column) {
    return(case_match(column,
            c("H-LC", "H-CE") ~ "H-CL",
            "S-CE" ~ "S-CL",
            "H-CTI" ~ "H-ST",
            "H-PS" ~ "H-MPPS",
            "S-FA" ~ "S-F",
            "S-GLT" ~ "S-GL",
            .default = column
        ))
}

manipulate <- function(df, Year) {
    return(df %>%
        mutate(Year = Year)  %>%
        separate(Com.1, into = c("Com.1", "Com.1.2", "Com.1.3", "Com.1.4"), sep = ";") %>%
        separate(Com.2, into = c("Com.2", "Com.2.2", "Com.2.3", "Com.2.4"), sep = ";") %>%
        mutate(Disposition = if_else(Disposition == "JRP", "PIL", Disposition)) %>%
        # filter(Disposition != "JRP") %>%
        mutate(
            Com.1 = sub(Com.1),
            Com.1.2 = sub(Com.1.2),
            Com.1.3 = sub(Com.1.3),
            Com.1.4 = sub(Com.1.4),
            Com.2 = sub(Com.2),
            Com.2.2 = sub(Com.2.2),
            Com.2.3 = sub(Com.2.3),
            Com.2.4 = sub(Com.2.4)
        )

    )
}

com_comparison <- function(df, list_of_committees) {
    in_pass <- 0
    out_pass <- 0
    in_fail <- 0
    out_fail <- 0
    for (i in seq_len(nrow(df))) {
        row <- df[i, ]

        if (is.na(row["Pass.Com.1"])) {
            if (row["Com.1"] %in% list_of_committees) {
                in_fail <- in_fail + 1
            } else {
                out_fail <- out_fail + 1
            }
            next
        }

        if (row["Pass.Com.1"] == 1) {
            if (row["Com.1"] %in% list_of_committees) {
                in_pass <- in_pass + 1
            } else {
                out_pass <- out_pass + 1
            }
            if (!is.na(row["Com.1.2"])) {
                if (row["Com.1.2"] %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }
            if (!is.na(row["Com.1.3"])) {
                if (row["Com.1.3"] %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }
            if (!is.na(row["Com.1.4"])) {
                if (row["Com.1.4"] %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }
        } else {
            # Doesn't pass committee 1
            if (is.na(row["Com.1.2"])) {
                if (row["Com.1"] %in% list_of_committees) {
                    in_fail  <- in_fail + 1
                    next
                } else {
                    out_fail <- out_fail + 1
                    next
                }
            } else {
                if (row["Com.1"] %in% list_of_committees) {
                    in_pass  <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
                if (is.na(row["Com.1.3"])) {
                    if (row["Com.1.2"] %in% list_of_committees) {
                        in_fail <- in_fail + 1
                        next
                    } else {
                        out_fail <- out_fail + 1
                        next
                    }
                } else {
                    if (row["Com.1.2"] %in% list_of_committees) {
                        in_pass <- in_pass + 1
                    } else {
                        out_pass <- out_pass + 1
                    }
                    if (is.na(row["Com.1.4"])) {
                        if (row["Com.1.3"] %in% list_of_committees) {
                            in_fail <- in_fail + 1
                            next
                        } else {
                            out_fail <- out_fail + 1
                            next
                        }
                    } else {
                        if (row["Com.1.3"] %in% list_of_committees) {
                            in_pass <- in_pass + 1
                        } else {
                            out_pass <- out_pass + 1
                        }
                        if (row["Com.1.4"] %in% list_of_committees) {
                            in_fail <- in_fail + 1
                            next
                        } else {
                            out_fail <- out_fail + 1
                            next
                        }
                    }
                }
            }
        }

        if (!is.na(row["Pass.Floor.1"])) {
            if (row["Pass.Floor.1"] == 0) {
                next
            }
        if (row["Pass.Com.2"] == 1) {
            if (row["Com.2"] %in% list_of_committees) {
                in_pass <- in_pass + 1
            } else {
                out_pass <- out_pass + 1
            }
            if (!is.na(row["Com.2.2"])) {
                if (row["Com.2.2"] %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }
            if (!is.na(row["Com.2.3"])) {
                if (row["Com.2.3"] %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }
            if (!is.na(row["Com.2.4"])) {
                if (row["Com.2.4"] %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }
        } else {
            # Doesn't pass committee 1
            if (is.na(row["Com.2.2"])) {
                if (row["Com.2"] %in% list_of_committees) {
                    in_fail  <- in_fail + 1
                        next
                    } else {
                        out_fail <- out_fail + 1
                        next
                    }
                } else {
                    if (row["Com.2"] %in% list_of_committees) {
                        in_pass  <- in_pass + 1
                    } else {
                        out_pass <- out_pass + 1
                    }
                    if (is.na(row["Com.2.3"])) {
                        if (row["Com.2.2"] %in% list_of_committees) {
                            in_fail <- in_fail + 1
                                next
                        } else {
                            out_fail <- out_fail + 1
                            next
                        }
                    } else {
                        if (row["Com.2.2"] %in% list_of_committees) {
                            in_pass <- in_pass + 1
                        } else {
                            out_pass <- out_pass + 1
                        }
                        if (is.na(row["Com.2.4"])) {
                            if (row["Com.2.3"] %in% list_of_committees) {
                                in_fail <- in_fail + 1
                            } else {
                                out_fail <- out_fail + 1
                            }
                        } else {
                            if (row["Com.2.3"] %in% list_of_committees) {
                                in_pass <- in_pass + 1
                            } else {
                                out_pass <- out_pass + 1
                            }
                            if (row["Com.2.4"] %in% list_of_committees) {
                                in_fail <- in_fail + 1
                                next
                            } else {
                                out_fail <- out_fail + 1
                                next
                            }
                        }
                    }
                }
            }
        }
    }

    tab <- as.table(rbind(c(in_pass, in_fail), c(out_pass, out_fail)))
    dimnames(tab) <- list(committee = c("in", "out"), pass_fail = c("pass", "fail"))
    # print(tab)
    # print(chisq.test(tab))
    return(c(in_pass, in_fail))
}
com_comparison_readable <- function(df, list_of_committees, print_table = FALSE) {
    in_pass <- 0
    out_pass <- 0
    in_fail <- 0
    out_fail <- 0
    df[df == ""] <- NA
    for (i in seq_len(nrow(df))) {
        # print(i)
        row <- df[i, ]
        if (row$Disposition %in% c("PIL", "DIF", "V")) {
            if (!is.na(row$Com.1)) {
                if (row$Com.1 %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }

            if (!is.na(row$Com.1.2)) {
                if (row$Com.1.2 %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }

            if (!is.na(row$Com.1.3)) {
                if (row$Com.1.3 %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }

            if (!is.na(row$Com.1.4)) {
                if (row$Com.1.4 %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }


            if (!is.na(row$Com.2)) {
                if (row$Com.2 %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }

            if (!is.na(row$Com.2.2)) {
                if (row$Com.2.2 %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }

            if (!is.na(row$Com.2.3)) {
                if (row$Com.2.3 %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }

            if (!is.na(row$Com.2.4)) {
                if (row$Com.2.4 %in% list_of_committees) {
                    in_pass <- in_pass + 1
                } else {
                    out_pass <- out_pass + 1
                }
            }
        } else {
            # Failed somewhere in the committees
            if (!is.na(row$Com.1)) {
                if (is.na(row$Com.1.2) && is.na(row$Com.2)) {
                    if (row$Com.1 %in% list_of_committees) {
                        in_fail <- in_fail + 1
                    } else {
                        out_fail <- out_fail + 1
                    }
                } else {
                    if (row$Com.1 %in% list_of_committees) {
                        in_pass <- in_pass + 1
                    } else {
                        out_pass <- out_pass + 1
                    }
                }
            }

            if (!is.na(row$Com.1.2)) {
                if (is.na(row$Com.1.3) && is.na(row$Com.2)) {
                    if (row$Com.1.2 %in% list_of_committees) {
                        in_fail <- in_fail + 1
                    } else {
                        out_fail <- out_fail + 1
                    }
                } else {
                    if (row$Com.1.2 %in% list_of_committees) {
                        in_pass <- in_pass + 1
                    } else {
                        out_pass <- out_pass + 1
                    }
                }
            }

            if (!is.na(row$Com.1.3)) {
                if (is.na(row$Com.1.4) && is.na(row$Com.2)) {
                    if (row$Com.1.3 %in% list_of_committees) {
                        in_fail <- in_fail + 1
                    } else {
                        out_fail <- out_fail + 1
                    }
                } else {
                    if (row$Com.1.3 %in% list_of_committees) {
                        in_pass <- in_pass + 1
                    } else {
                        out_pass <- out_pass + 1
                    }
                }
            }

            if (!is.na(row$Com.1.4)) {
                if (is.na(row$Com.2)) {
                    if (row$Com.1.4 %in% list_of_committees) {
                        in_fail <- in_fail + 1
                    } else {
                        out_fail <- out_fail + 1
                    }
                } else {
                    if (row$Com.1.4 %in% list_of_committees) {
                        in_pass <- in_pass + 1
                    } else {
                        out_pass <- out_pass + 1
                    }
                }
            }

            if (!is.na(row$Com.2)) {
                if (is.na(row$Com.2.2)) {
                    if (row$Com.2 %in% list_of_committees) {
                        in_fail <- in_fail + 1
                    } else {
                        out_fail <- out_fail + 1
                    }
                } else {
                    if (row$Com.2 %in% list_of_committees) {
                        in_pass <- in_pass + 1
                    } else {
                        out_pass <- out_pass + 1
                    }
                }
            }

            if (!is.na(row$Com.2.2)) {
                if (is.na(row$Com.2.3)) {
                    if (row$Com.2.2 %in% list_of_committees) {
                        in_fail <- in_fail + 1
                    } else {
                        out_fail <- out_fail + 1
                    }
                } else {
                    if (row$Com.2.2 %in% list_of_committees) {
                        in_pass <- in_pass + 1
                    } else {
                        out_pass <- out_pass + 1
                    }
                }
            }
            if (!is.na(row$Com.2.3)) {
                if (is.na(row$Com.2.4)) {
                    if (row$Com.2.3 %in% list_of_committees) {
                        in_fail <- in_fail + 1
                    } else {
                        out_fail <- out_fail + 1
                    }
                } else {
                    if (row$Com.2.3 %in% list_of_committees) {
                        in_pass <- in_pass + 1
                    } else {
                        out_pass <- out_pass + 1
                    }
                }
            }

            if (!is.na(row$Com.2.4)) {
                if (row$Com.2.4 %in% list_of_committees) {
                    in_fail <- in_fail + 1
                } else {
                    out_fail <- out_fail + 1
                }
            }
        }
    }

    tab <- as.table(rbind(c(in_pass, in_fail), c(out_pass, out_fail)))
    dimnames(tab) <- list(committee = c("in", "out"), pass_fail = c("pass", "fail"))
    if (print_table) {
        print(tab)
        print(chisq.test(tab))
    }
    return(c(in_pass, in_fail))
}

party_comparison <- function(df, print_table = FALSE) {
    tab_party <- matrix(c(
        "R", "R",
        "R", "R",
        "R", "R",
        "R", "R",
        "R", "R",
        "D", "D",
        "D", "D",
        "R", "D",
        "R", "D"
    ), ncol=2, byrow=TRUE)
    colnames(tab_party) <- c("H", "S")
    rownames(tab_party) <- seq(2015, 2023)
    tab_party <- data.frame(tab_party)
    tab <- as.table(rbind(c(0, 0), c(0, 0)))
    dimnames(tab) <- list(party = c("D", "R"), pass_fail = c("pass", "fail"))
    for (i in seq_len(nrow(df))) {
        row <- df[i, ]
        c1 <- substring(row["Com.1"], 1, 1)
        p1 <- tab_party[as.character(row["Year"]), c1]
        c2 <- substring(row["Com.2"], 1, 1)
        # If failed committee 1
        if (c2 == "") {
            tab[p1, "fail"] <- tab[p1, "fail"] + 1
        } else if (!is.na(row["Pass.Floor.2"]) && row["Pass.Floor.2"] == 1) {
            # If passed both committees
            p2 <- tab_party[as.character(row["Year"]), c2]
            tab[p1, "pass"] <- tab[p1, "pass"] + 1
            tab[p2, "pass"] <- tab[p2, "pass"] + 1
        } else {
            # Only passed committee 1
            p2 <- tab_party[as.character(row["Year"]), c2]
            tab[p1, "pass"] <- tab[p1, "pass"] + 1
            tab[p2, "fail"] <- tab[p2, "fail"] + 1
        }
    }

    if (print_table) {
        print(tab)
        print(chisq.test(tab))
    }
    return(tab)
}
```

## Data Import

Now, import the bill data.
The following are made:

* `csvxx` - Data for the year 20xx
* `csv_total` - Combines all `csvxx` into one, combining `DiH` and `DiS` into `DiF`
* `climate` - Addition to `csv_total`, adding fate (`DIC`, `PIL`, `DIO`), and dem (whether the year is democratic trifecta)
* `df` - A separate dataframe that keeps track of how many bills passed and failed in each committee per year
* `cl` - Removes neutral bills, had `DIC`, `PIL`, `DIO`

```{r Data Import, echo=FALSE}
overall <- read.csv("data/overall.csv")
csv15 <- filter(overall, Session == 2015) %>% col_care() %>% manipulate(Year = 2015) %>% add_identifiers() %>% rename(prog = SC.Position)
csv16 <- filter(overall, Session == 2016) %>% col_care() %>% manipulate(Year = 2016) %>% add_identifiers() %>% rename(prog = SC.Position)
csv17 <- filter(overall, Session == 2017) %>% col_care() %>% manipulate(Year = 2017) %>% add_identifiers() %>% rename(prog = SC.Position)
csv18 <- filter(overall, Session == 2018) %>% col_care() %>% manipulate(Year = 2018) %>% add_identifiers() %>% rename(prog = SC.Position)
csv19 <- filter(overall, Session == 2019) %>% col_care() %>% manipulate(Year = 2019) %>% add_identifiers() %>% rename(prog = SC.Position)
csv20 <- filter(overall, Session == 2020) %>% col_care() %>% manipulate(Year = 2020) %>% add_identifiers() %>% rename(prog = SC.Position)
csv21 <- filter(overall, Session == 2021) %>% col_care() %>% manipulate(Year = 2021) %>% add_identifiers() %>% rename(prog = SC.Position)
csv22 <- filter(overall, Session == 2022) %>% col_care() %>% manipulate(Year = 2022) %>% add_identifiers() %>% rename(prog = SC.Position)
csv23 <- filter(overall, Session == 2023) %>% col_care() %>% manipulate(Year = 2023) %>% add_identifiers() %>% rename(prog = SC.Position)

csv_total <- rbind(csv15, csv16, csv17, csv18, csv19, csv20, csv21, csv22, csv23) %>%
    mutate(Disposition = if_else(Disposition == "", "INC", str_to_upper(Disposition))) %>%
    mutate(Disposition = if_else(Disposition %in% c("DIH", "DIS"), "DIF", Disposition))


climate <- data.frame(fate = csv_total$Disposition[csv_total$Disposition != "JRP"], bill = "Climate", year = csv_total$Year[csv_total$Disposition != "JRP"], prog = csv_total$prog[csv_total$Disposition != "JRP"])  %>% mutate(fate = if_else(fate %in% c("DIC", "PIL"), fate, "DIO")) %>%
    mutate(dem = year %in% c(2020, 2021))

cl <- filter(csv_total, Disposition != "JRP") %>%
    filter(prog != 0) %>%
    # mutate(prog = prog == 1) %>%
    mutate(Disposition = if_else(Disposition %in% c("DIC", "PIL"), Disposition, "DIO"))


df <- data.frame(matrix(ncol = 4, nrow = 200))
colnames(df) <- c("year", "committee", "pass", "fail")
i <- 1
for (year in 2015:2023) {
    # print(year)
    subset <- filter(csv_total, Year == year)

    set_committees <- na.omit(unique(c(csv_total$Com.1, csv_total$Com.1.2, csv_total$Com.1.3, csv_total$Com.1.4, csv_total$Com.2, csv_total$Com.2.2, csv_total$Com.2.3, csv_total$Com.2.4))) %>% .[. != ""]
    for (com in set_committees) {
        result <- com_comparison_readable(subset, com)
        df[i, ] <- c(year, com, result[1], result[2])
        # print(result)
        i <- i + 1
    }
}
df$year <- as.numeric(as.character(df$year))
df$pass <- as.numeric(as.character(df$pass))
df$fail <- as.numeric(as.character(df$fail))
# write.csv(df, "climate_com_pass_fail_year.csv", row.names = FALSE)
```

# Analysis

## General Analysis

Number of bills between 2015 and 2023

```{r}
nrow(csv_total)
```

Count how many bills passed through each committee overall

```{r}
mutate(df, total = pass + fail) %>%
    group_by(committee) %>%
    summarize(pass = sum(pass), fail = sum(fail), total = sum(total)) %>%
    print(n = 50)
```

Over 100 bills?

```{r}
mutate(df, total = pass + fail) %>%
    group_by(committee) %>%
    summarize(pass = sum(pass), fail = sum(fail), total = sum(total)) %>%
    filter(total >= 100)
```

How many bills per year?

```{r}
group_by(csv_total, Year) %>%
    summarize(n = n())
```

How many progressive bills?

```{r}
filter(csv_total, Pos == "Supported") %>%
    group_by(Year) %>%
    summarize(n = n())
```

What percent of legislation passes into law per year?

```{r}
group_by(csv_total, Year) %>%
    summarize(rate = mean(Disposition == "PIL"))
```

How does democratic trifecta affect passage rates?

```{r}
group_by(climate, dem) %>%
    summarize(percent = mean(fate == "PIL"))
(tab <- table(climate$dem, climate$fate))
chisq.test(tab)
```

How about for just progressive legislation?

```{r}
progressive <- filter(climate, prog == 1)
group_by(progressive, dem) %>%
    summarize(percent = mean(fate == "PIL"))
(tab <- table(progressive$dem, progressive$fate))
chisq.test(tab)
```

How about non-progressive?

```{r}
regressive <- filter(climate, prog != 1)
group_by(regressive, dem) %>%
    summarize(percent = mean(fate == "PIL"))
(tab <- table(regressive$dem, regressive$fate))
chisq.test(tab)
```

How does whether a bill is progressive change things?

```{r}
group_by(climate, prog) %>%
    summarize(percent = mean(fate == "PIL"))

(tab <- table(climate$prog, climate$fate == "PIL"))
chisq.test(tab)
```

Remove neutral bills because those have high passage rate and are few

```{r}
not_neutral <- filter(climate, prog != 0)
group_by(not_neutral, prog) %>%
    summarize(percent = mean(fate == "PIL"))

(tab <- table(not_neutral$prog, not_neutral$fate))
chisq.test(tab)

(tab <- table(not_neutral$prog, not_neutral$fate == "PIL"))
chisq.test(tab)
```

What about year by year?

```{r}
for (y in 2015:2023) {
    print(paste("Starting year", y))
    subset <- filter(not_neutral, year == y)
    print(tab <- table(subset$prog, subset$fate))
    print(chisq.test(tab))
}
```

Where do bills die?

```{r}
dead <- filter(climate, fate != "PIL")
print(table(dead$fate))
print(table(dead$fate) / nrow(dead))
```

How about for different progressiveness?

```{r}
tab <- table(dead$fate, dead$prog)
print(tab)
print(sweep(tab, 2, colSums(tab), "/"))
```

What part of bills that died in committee died in their first committee?

```{r}
dead_in_committee <- filter(csv_total, Disposition == "DIC")
dead_in_first <- filter(dead_in_committee, Com.2 == "")
nrow(dead_in_first) / nrow(dead_in_committee)
```

## Analysis of CL and ACNR

Bills overall

```{r}
com_comparison_readable(csv_total, "H-CL", TRUE)
com_comparison_readable(csv_total, "S-CL", TRUE)
com_comparison_readable(csv_total, "H-ACNR", TRUE)
com_comparison_readable(csv_total, "S-ACNR", TRUE)
```

Only progressive bills

```{r}
progressive <- filter(csv_total, Pos == "Supported")
com_comparison_readable(progressive, "H-CL", TRUE)
com_comparison_readable(progressive, "S-CL", TRUE)
com_comparison_readable(progressive, "H-ACNR", TRUE)
com_comparison_readable(progressive, "S-ACNR", TRUE)
```

Only regressive bills

```{r}
regressive <- filter(csv_total, Pos == "Opposed")
com_comparison_readable(regressive, "H-CL", TRUE)
com_comparison_readable(regressive, "S-CL", TRUE)
com_comparison_readable(regressive, "H-ACNR", TRUE)
com_comparison_readable(regressive, "S-ACNR", TRUE)
```

Let's look at how party affects everything

```{r}
party_comparison(csv_total, TRUE)
```

And progressive bills

```{r}
t <- party_comparison(progressive, TRUE)
```

And finally regressive bills

```{r}
t <- party_comparison(regressive, TRUE)
```

What about when there is a split house?

```{r}
filter(csv_total, Year %in% c(2022, 2023)) %>%
    party_comparison(TRUE)
filter(progressive, Year %in% c(2022, 2023)) %>%
    party_comparison(TRUE)
t <- filter(regressive, Year %in% c(2022, 2023)) %>%
    party_comparison(TRUE)
```
